Measuring the Type Ia supernova delay time distribution via host galaxy rates in the Dark Energy Survey

### 2021-01-13 draft.tex

@@

%%%% INCOMPLETENESS %%%%
\section{Incompleteness corrections}
\label{sec:incompleteness}
Although Eq. \ref{eq:rate1} serves as an approximation to the SN rate per galaxy per year, both the SN and field galaxy samples introduced in Sections \ref{subsec:host_sample} and \ref{subsec:field_sample} are affected by incompleteness, which is likely to be the dominant systematic effect in the analysis. In this section we describe our method of correcting for various sources incompleteness in the data.

@@

\subsubsection{Apparent magnitude limits \label{subsubsec:mag_lims}}
As we use photometric, rather than spectroscopic, redshifts for the field galaxies, they do not suffer from spectroscopic incompleteness as the SN hosts do. Instead, the inclusion of any given galaxy in the survey area in the sample is determined simply by whether it is detected above a prescribed threshold in signal-to-noise ratio, i.e. the sample is magnitude limited. To determine the apparent magnitude limit in each of the optical bands we employ the method of \citet{Johnston2007,Teodoro2010,Johnston2012} (hereafter Completeness I, II, III respectively). Test statistics $T_C$ and $T_V$ are computed based on absolute magnitude and distance modulus by calculating the rank of each galaxy's absolute magnitude (distance modulus) when compared to all other galaxies in a survey within a certain slice of distance modulus (absolute magnitude). For a galaxy of apparent magnitude $m < m_{\mathrm{lim, trial}}$ observed in a survey complete to magnitude $m_{\mathrm{lim, true}}$ where  $m_{\mathrm{lim, trial}} \leq m_{\mathrm{lim, true}}$, the expectation value of the rank is 0.5. However, if a trial limiting magnitude $m_{\mathrm{lim, trial}} \geq m_{\mathrm{lim, true}}$, there will be a lack of observed faint objects, such that the expectation value of the rank drops. Completeness I, II, and III show that for surveys with sharp, well defined magnitude limits the test statistics $T_C$ and $T_V$ drop sharply. The DES deep stacks present a more challenging case. They cover a wide area and are compiled from tens of separate CCDs, each with its own detection efficiency. Moreover, galaxy redshifts have been estimated using photometry only, and $K$-corrections performed using an imperfect template fit, potentially leading to large systematic and statistical fluctuations in absolute magnitudes and distance moduli. To reduce the inhomogeneity while still including  large sample of objects, we split the sample into the deep (X3) and shallow (E2) fields and calculate limiting magnitudes separately. Figs. \ref{fig:completeness_X3, fig:completenessE2} show the values of $T_C$ and $T_V$ for the two fields respectively. The values of the test statistics increase with $m_{\mathrm{lim, trial}}$ until a peak, before decreasing to stable values at magnitudes far beyond the limit of the survey. The shape at brighter $m_{\mathrm{lim, trial}}$ is likely caused by incompleteness at the bright end: due to the small sky area, we simply do not probe enough volume to sample the bright end of the galaxy luminosity function well enough for the statistics to be robust. To approximate an efficiency function, we fit the peak of the statistics with a polynomial function, and then normalise by the maximum and minimum values. We then interpolate between the peak and the faint-magnitude floor to find the 50\% completeness limit at the point where the normalised value of the test statistic is 0.5. These values are consistent with the value at which 50\% of the true sources are detected but have the advantage of being derived from the data without a simulation that is based on assumptions and thus susceptible to bias.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/completeness_combined.png}
    \caption{Measuring the completeness of the field galaxy sample. \textit{Upper}: the $T_{\mathrm{C}}$ statistic as a function of apparent magnitude; \textit{middle} as per \textit{upper}, but for $T_{\mathrm{V}}$; \textit{lower}: the combined, normalised completeness in each band.  
    \label{fig:completeness}}
\end{figure}

@@

%%%% INCOMPLETENESS %%%%

\section{Incompleteness corrections}
\label{sec:incompleteness}
Although Eq. \ref{eq:rate1} serves as an approximation to the SN rate per galaxy per year, both the SN and field galaxy samples introduced in Sections \ref{subsec:host_sample} and \ref{subsec:field_sample} are affected by incompleteness, which is likely to be the dominant systematic effect in the analysis. In this section we describe our method of correcting for various sources incompleteness in the data.

@@

\subsubsection{Apparent magnitude limits \label{subsubsec:mag_lims}}
As we use photometric, rather than spectroscopic, redshifts for the field galaxies, they do not suffer from spectroscopic incompleteness as the SN hosts do. Instead, the inclusion of any given galaxy in the survey area in the sample is determined simply by whether it is detected above a prescribed threshold in signal-to-noise ratio, i.e. the sample is magnitude limited. To determine the apparent magnitude limit in each of the optical bands we employ the method of \citet{Johnston2007,Teodoro2010,Johnston2012} (hereafter Completeness I, II, III respectively). Test statistics $T_{\mathrm{C}}$ and $T_{\mathrm{V}}$ are computed based on absolute magnitude and distance modulus by calculating the rank of each galaxy's absolute magnitude (distance modulus) when compared to all other galaxies in a survey within a certain slice of distance modulus (absolute magnitude). For a galaxy of apparent magnitude $m < m_{\mathrm{lim, trial}}$ observed in a survey complete to magnitude $m_{\mathrm{lim, true}}$ where  $m_{\mathrm{lim, trial}} \leq m_{\mathrm{lim, true}}$, the expectation value of the rank is 0.5. However, if a trial limiting magnitude $m_{\mathrm{lim, trial}} \geq m_{\mathrm{lim, true}}$, there will be a lack of observed faint objects, such that the expectation value of the rank drops. Completeness I, II, and III show that for surveys with sharp, well defined magnitude limits the test statistics $T_{\mathrm{C}}$ and $T_{\mathrm{V}}$ drop sharply. The DES deep stacks present a more challenging case. They cover a wide area and are compiled from tens of separate CCDs, each with its own detection efficiency. Moreover, galaxy redshifts have been estimated using photometry only, and $K$-corrections performed using an imperfect template fit, potentially leading to large systematic and statistical fluctuations in absolute magnitudes and distance moduli. To reduce the inhomogeneity while still including  large sample of objects, we split the sample into the deep (X3) and shallow (E2) fields and calculate limiting magnitudes separately. Fig. \ref{fig:completeness} shows the values of $T_{\mathrm{C}}$ and $T_{\mathrm{V}}$. The values of the test statistics increase with $m_{\mathrm{lim, trial}}$ until a peak, before decreasing to stable values at magnitudes far beyond the limit of the survey. The shape at brighter $m_{\mathrm{lim, trial}}$ is likely caused by incompleteness at the bright end: due to the small sky area, we simply do not probe enough volume to sample the bright end of the galaxy luminosity function well enough for the statistics to be robust. To approximate an efficiency function, we fit the peak of the statistics with a polynomial function, and then normalise by the maximum and minimum values. We then interpolate between the peak and the faint-magnitude floor to find the 50\% completeness limit at the point where the normalised value of the test statistic is 0.5. These values are consistent with the value at which 50\% of the true sources are detected but have the advantage of being derived from the data without a simulation that is based on assumptions and thus susceptible to bias.
